From Real-time Attention Assessment to ``With-me-ness'' in Human-Robot Interaction

### 2015-09-22 paper.tex
WIP on the discussion

\paragraph{Practicality and Relevance of 2D Methods for Head Pose Estimation}

Furthermore, although it has not been tested in a real scenario with several
subjects at the same time, the approach also allows to handle such situations
too.

@@

Regarding to the approach used, some limitations need to be considered. The most
relevant one, the fact that the positions of the VFoA need to be known in
advance and defined as static \textit{tf} frames in the system. Moreover, it is
important to notice that eye gaze information is neglected which induces a
certain amount of error, as well as the noise induced by the 3D pose estimation
computed from face features.

Besides, our approach does not account for visual occlusions when computing the
focus of attention: any frame within the cone of the visual field of attention
is considered as \emph{in focus}. In reality, some of the frames may be visually
occluded and thus not seen by the subject. Computing actual visibility requires
more general situation assessment capabilities, and in particular (approximate)
3D models of the surrounding objects so that visibility testing may be
performed, as done in~\cite{sisbot2011situation} for instance. Our experimental
setting did not lead to any specific visual occlusions, and we did not developed
this capability.

Compared to intrusive hardware like head mounted
displays, eye-trackers or wearables for instance, this approach has a clear
advantage of simplicity and integration with robotic systems, but also becomes
accurate enough for general purposes in face-to-face interactions. In addition,
it allows us to model complex environments, handling different number and
positions of participants in the scene.

@@

>>>>

\paragraph{Head Pose to Assess Attention: is it Relevant?}

@@

We already stated the main limitations of our approach to estimation of the
focus of attention: eye gaze information is neglected and we do not perform
actual visibility check of the in focus objects (we simply approximate them to
their origins, ignoring possible occlusions).

The first issue is shared with most of the other vision (2D or 3D) or motion
capture techniques for gaze estimation. In that regard, our approach does not
perform worst or better than others. In fact, based on our results, relying
purely on head pose estimation to estimate gaze direction leads to real-world
results that are worth being considered and used: they may not match manual
annotations, but they are definitely a valuable \emph{in-the-moment} input for
the robot, and we hope to see more works relying on those to build adaptive
robot controllers in the coming years.

@@

Besides, our approach relies on a simple, non-intrusive sensor (a RGB camera on
the robot) and an open-source, fast pose estimation algorithm\fixme{test on Nao!
how fast is it actually?}: we also hope that this may contribute to a widespread
adoption of such a technique on a range of robots, including the common {\sc
Nao} platform.
